---
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "../output_word") })
title: "Crab molts for exoskeleton pilot 2021"
output:
  officedown::rdocx_document:
    reference_docx: word_style_ref_01.docx
editor_options:
  chunk_output_type: console
---

```{r knitter-options, include = FALSE}
# chunk header include = FALSE supresses code in the Word output
# chunk options suppress warnings, messages and output in the Word output
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=FALSE)
```

```{r libraries}
library(here)
library(tidyverse)
library(janitor)
library(officedown)
library(metacsv)
library(todor)
```

## Rosario Zoea

There two data files (crab care and preservation).

```{r rosaio-zoea-read-and-clean-collect-care}

#read the status key from the crab care file (first row of file)
d_rosario_care_key <- read.csv(here("data_raw", "crab_zoea_exoskel_rosario_carechart_2021.csv"),header = FALSE, nrows = 1)[,1] 

#read in the crab care data file data
d_rosario_care_raw <- read_csv(here("data_raw", "crab_zoea_exoskel_rosario_carechart_2021.csv"), skip = 3) 

#funtion to create new variable names for the "By" columns (person who entered
#data) that they includes the date
new_name_care <- function(old_names){
  new_names <- old_names
  for(i in 1:length(old_names)){
    if(str_starts(old_names[i], "By")){
      new_names[i] <- paste("entered_by_", old_names[i+1], sep = "")
    }
  }
  return(new_names)
} 
#rename care file variables
names(d_rosario_care_raw) <- new_name_care(names(d_rosario_care_raw))


# read clean up the crab care data frame and add a "fate" column to indicate id crab died or molted.
# note that the case_when should work, but there is a bug in if_any (https://github.com/tidyverse/dplyr/issues/5782#)
d_rosario_care <- d_rosario_care_raw %>%
  filter(if_any(everything(), ~ !is.na(.))) %>%
  mutate(crabID = paste(crabID, startLoc, sep = "_")) %>%
  select_if(function(x){!all(is.na(x))}) %>%
  # mutate(fate = case_when(if_any(everything(), ~ . == "D") ~ "dead",
  #                         if_any(everything(), ~ . == "L") ~ "lost",
  #                         if_any(everything(), ~ . == "M") ~ "molt",
  #                         if_any(everything(), ~ . == "R") ~ "removed",
  #                         TRUE ~ "whatever")) %>%
  mutate(fate = NA_character_, 
    fate = if_else(if_any(everything(), ~ . == "D"),  "dead", fate),
    fate = if_else(if_any(everything(), ~ . == "M"),  "molted", fate),
    fate = if_else(if_any(everything(), ~ . == "L"),  "lost", fate),
    fate = if_else(if_any(everything(), ~ . == "R"),  "removed", fate),
    fate = if_else(is.na(fate), "alive_unmolted", fate)) %>%
  {.}

# Read in the collection and experiment start data
d_rosario_collect <- read_csv(here("data_raw", "crab_zoea_exoskel_rosario_collection_2021.csv"), skip = 2) %>%
  mutate(crabID = paste(crabID, startLoc, sep = "_")) %>%
  {.}

#merge the crab care and collection files
d_rosario <- full_join(d_rosario_care, d_rosario_collect, by = "crabID")


```

```{r rosaio-zoea-read-and-clean-preserve}

# read and recode the rosario perservation file
d_rosario_preservation <- read_csv(here("data_raw", "brachyura_zoe_preservationLog_popvials_DANIELLE'S_NOTES.csv")) %>%
  mutate(preserve_stage = case_when(preservationDate == "6/9/21" & lifeStage == "zoea" ~ "zoea_start",
                                preservationDate == "6/23/21"  & lifeStage == "zoea" ~ "zoea_end",
                                lifeStage == "zoea" ~ "zoea_live_post_molt",
                                lifeStage == "megalopae" ~ "meg_live_post_molt",
                                lifeStage == "Molt" | lifeStage == "zoeaMOLT" ~ "discard_molt",
                                TRUE ~ NA_character_)) %>%
  mutate(crabType = case_when(species == "brachyura_ssp." ~ "Pea",
                              species == "C.magister" ~ "Dungeness",
                              TRUE ~ NA_character_)) %>%
  {.}
```

A few question about the data summary below. In Table \@ref(tab:preserve-summary), is the assumption that crabs labeled "Instar" molted to a later stage zoea? In Table \@ref(tab:preserve-summary), a total of 20 discarded molts were preserved but a total only 18 live zoea were preserved (zoea + megalopae) - what happened to the other live crab? In table \@ref(tab:fate-summary), there were 18 molted pea crab and 2 dungeness. However, in Table \@ref(tab:preserve-summary), 20 molts were preserved that were all labeled pea crab. There were also no dungeness preserved as live post-molt

```{r rosario-zoea-fate-summary-tab, tab.cap = "Summary of the fate of zoea from Rosario (round 1). During the experiment, zoea could 1) die before, 2) molt, in which case both the new stage zoea or megalope and the discarded exoskeleton were preserved in seperate vials, or 3) the zoea could remain alivee and unmolted at the end of the experinment , in which case they were preserved. CrabType is presumed (crabs have not been IDed to species).", tab.id = "fate-summary"}

# FIXME Danielle and Paul get different output - should be same code and data- why???

d_rosario_care %>%
  tabyl(crabType, fate) %>%
  adorn_totals(where = c("row", "col")) %>%
  as.data.frame()
```

```{r rosario-zoea-preserve-summary, tab.cap = "Preserve sample summary based on preservation csv file. Zoea_start are zoea preserved from the coolers at the beginnig of the experiment - they were never in chambers. Zoea_live and meg_live are crabs preseverd after they molted. It was assumed that the crabs marked Instar in the oringal preservation file molted to a later stage zoea and crabs marked Megalopae molted to megalopae. Discard_molts were the discarded exoskeletons after molting (marked either Molt or zoeaMOLT in the data sheet). Zoea_end are crabs that were alive at the end of the experiment but never molted.", tab.id = "preserve-summary"}
s_rosario_preservation <- d_rosario_preservation %>%
  group_by(crabType,preserve_stage) %>%
  summarise(sum(n_individuals.per.Vial)) %>%
  arrange(crabType, match(preserve_stage, 
                          c("zoea_start", "zoea_live", "meg_live", 
                            "discard_molt", "zoea_end"))) %>%
  rename("total" = "sum(n_individuals.per.Vial)") %>%
  as.data.frame()
s_rosario_preservation
```

```{r muk-meg-read-and-clean}

d_muk_meg_care_raw <- read_csv(here("data_raw", 
                                    "dng_meg_exoskel_muk_net_carechart_2021.csv"), skip = 1)

#rename care file variables
names(d_muk_meg_care_raw) <- new_name_care(names(d_muk_meg_care_raw))


# read clean up the crab care data frame and add a "fate" column to indicate id
# crab died or molted and "preserved" column to indicate which crab were
# preserved
d_muk_meg_care <- d_muk_meg_care_raw %>%
  filter(if_any(everything(), ~ !is.na(.))) %>%
  filter(crabID != "N/A") %>%
  select_if(function(x){!all(is.na(x))}) %>%
  select(crabID, starts_with(c("6", "7"))) %>%
  mutate(fate = NA_character_, 
    fate = if_else(if_any(everything(), ~ . %in% c("M, R", 
                                                   "Molt, Remove", 
                                                   "Molt, R (molt dissolve)", 
                                                   "Molt, Removed",
                                                   "Molt, R-(molt dissolved)")),"molted", fate),
    fate = if_else(`6/28` == "meg" & `6/30` == "J", "molted", fate),
    fate = if_else(if_any(everything(), ~ . == "D"),  "dead", fate),
    fate = if_else(if_any(everything(), ~ . == "R"),  "removed", fate),
    fate = if_else(is.na(fate), "alive_unmolted", fate)) %>%
  mutate(preserved_juv = if_else(fate == "molted",  TRUE, FALSE)) %>%
  mutate(preserved_discard = if_else(fate == "molted",  TRUE, FALSE),
         preserved_discard = if_else(if_any(everything(), ~ . %in% c( "Molt, R (molt dissolve)", 
                                                  "Molt, R-(molt dissolved)")), FALSE, preserved_discard)) %>%
  mutate(preserved_meg_end = if_else(`7/14` == "R, removed &preserved",  TRUE, FALSE)) %>%
  mutate(preserved = case_when(preserved_juv & !preserved_discard ~ "juv_only",
                               preserved_juv & preserved_discard ~ "juv_and_discard",
                               preserved_meg_end ~ "meg_end",
                               TRUE ~ "nothing")) %>%
  {.}

write_csv(d_muk_meg_care, here("data", "d_muk_meg_care_clean.csv"))

```

```{r muk-meg-care-summary}
#summary table based on care file
d_muk_meg_care %>%
  tabyl(fate, preserved) %>%
  adorn_totals(where = c("row", "col"))
```

```{r}
# CHANGED Danielle code edited by Paul on 2021.11.23

#summary table based on transfer sheet file

#Danielle original
# d_muk_meg_NEW_care <- read.csv(here("data", "WellEventData.csv"))
# d_muk_meg_NEW_care$molt <- as.integer(d_muk_meg_NEW_care$molt)
# s_muk_meg_NEW_care <- d_muk_meg_NEW_care %>%
#   group_by(crabID %>%
#   summarise(sum(molt)) %>%
#     as.data.frame())
# 
# s_muk_meg_NEW_care

#Paul revision
# Turn out you don't have to explictly convert logical variable (molt) to 0/1,
# if you do math on logical variable converson happens automatically
#problem in origal code was not closing the parthensis in the group_by function call
# you closed it after the as.data.frame call which is the wrong place.
# group_by() is sort of invisible - it changes how functions that come after it, like summarize work, 
# but you don't really see any difference in the dataframe running it by itself except for a hidden attribute.
d_muk_meg_NEW_care <- read_csv(here("data", "WellEventData.csv"))
s_muk_meg_NEW_care <- d_muk_meg_NEW_care %>%
  group_by(crabID) %>%
  summarise(molted = sum(molt)) %>%
  tabyl(molted) %>%
  adorn_totals() %>%
  as.data.frame() %>%
  {.}

s_muk_meg_NEW_care

```

```{r muk-meg-preserve-read-summary}
#read muk meg preservation file
d_muk_meg_preservation <- read_csv(here("data_raw", "dng_meg_exoskel_muk_net_preservation_2021.csv")) %>%
  filter(!is.na(lifeStage))

#summarize life stages presereved
s_muk_meg_preservation <- d_muk_meg_preservation %>%
  tabyl(lifeStage) %>%
  adorn_totals()
s_muk_meg_preservation
```

```{r}
d_coronet_care_raw <- read_csv(here("data_raw", "dng_meg_exoskel_coronet_carechart_2021.csv"), skip = 3)
#rename care file variables
names(d_coronet_care_raw) <- new_name_care(names(d_coronet_care_raw))


d_coronet_care <- d_coronet_care_raw %>%
  filter(if_any(everything(), ~ !is.na(.))) %>%
  filter(str_starts(crabID, "CH")) %>%
   select_if(function(x){!all(is.na(x))}) %>%
   select(crabID, starts_with(c("6", "7"))) %>%
  mutate(fate = NA_character_,
    fate = if_else(if_any(everything(), ~ str_detect(., "J1_frozen")), "j1_frozen", fate),
    fate = if_else(if_any(everything(), ~ str_detect(., fixed("J1_EtOH", ignore_case = TRUE))),  "j1_etoh", fate),
    fate = if_else(if_any(everything(), ~ . == "D"),  "dead", fate),
    fate = if_else(is.na(fate), "alive_unmolted", fate)) %>%
  mutate(preserved_juv = case_when(str_detect(fate, "frozen") ~ "frozen",
                                   str_detect(fate, "etoh") ~ "etoh",
                                   TRUE ~ NA_character_)) %>%
  mutate(preserved_discard = if_else(!is.na(preserved_juv) & 
                                       !str_detect(`7/12`, "dissolved"),
                                     preserved_juv, NA_character_)) %>%
  mutate(preserved_meg_end = if_else(fate == "alive_unmolted", TRUE, FALSE)) %>%
  mutate(preserved = case_when(!is.na(preserved_juv) & !is.na(preserved_discard) ~
                               paste("both_juv_and_discard_", preserved_juv, sep = "" ),
                               !is.na(preserved_juv) & is.na(preserved_discard) ~
                                paste("juv_only_", preserved_juv, sep = "" ),
                               preserved_meg_end ~"meg_end",
                               TRUE ~ "nothing")) %>%
  {.}
```

```{r}
d_coronet_care %>%
  tabyl(fate, preserved) %>%
   adorn_totals(where = c("row", "col")) %>%
  as.data.frame()

```

```{r}
d_coronet_preservation <- read_meta(here("data_raw", "dng_meg_exoskel_coronet_preservation_2021_with_meta.csv"))$data

s_coronet_preservation <- d_coronet_preservation %>%
  tabyl(preservationMaterial, vial) %>%
  adorn_totals(where = c("row", "col")) %>%
  as.data.frame()
s_coronet_preservation
```

```{r}
s_rosario_preservation
s_muk_meg_preservation
s_coronet_preservation
```
